### 2024-09-25 paper/paper.qmd
Update paper.qmd for the data part

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Jiwon Choi
thanks: "Code and data are available at: https://github.com/jwonc4602/Analysis-of-Motor-Vehicle-Collision-Causes."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#### Workspace set-up ####

## Load packages
library(tidyverse)
library(knitr)
library(here)
library(dplyr)
library(ggplot2)
library(readr)
library(gridExtra)
library(grid)

#read in cleaned data #
#total_cleaned_data
cleaned_data <- read.csv(here::here("data/analysis_data/total_cleaned_data.csv"))
show_col_types = FALSE

# human factors data
human_factors_data <- read.csv(here::here("data/analysis_data/human_factors_data.csv"))
show_col_types = FALSE

# environmental factors data
environmental_factors_data <- read.csv(here::here("data/analysis_data/environmental_factors_data.csv"))
show_col_types = FALSE

# result data
result_data <- read.csv(here::here("data/analysis_data/result_data.csv"))
show_col_types = FALSE
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Data Source

This research is based on a dataset published by Toronto Police Services, which is accessible through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto]. The specific dataset includes all traffic collisions events in Toronto where a person was either Killed or Seriously Injured (KSI) since 2006.

## Data Collection

The methodology for gathering data is structured to ensure a detailed and current perspective of shelter and overnight service. The data was compiled and examined using the R statistical programming software [@r], supplemented by various tools such as `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `knitr` [@rknitr], and `here` [@rHere].

This traffic accident data consists of 18,957 entries from a citywide accident survey, capturing detailed information about various factors involved in traffic incidents. The raw dataset includes 50 variables, ranging from geographic data, such as the street where the accident occurred, to behavioral data, such as driver condition and alcohol involvement. This dataset records accidents over multiple years and provides insights into environmental conditions, road types, and the actions of drivers and pedestrians at the time of the incidents. The goal of this dataset is to provide a comprehensive view of traffic accidents for research and policy analysis related to urban traffic safety and accident prevention.

The cleaned data is divided into three datasets, and the first one focuses on human factors related to traffic collisions, specifically the behaviors and conditions of drivers that may have contributed to accidents. The four main categories—'speeding', 'aggressive_and_disturbed_driving', 'alcohol', and 'disability'—were selected because these are well-known, critical factors that significantly increase the risk of traffic accidents. Each factor is evaluated as either "None" or "Yes", indicating the absence or presence of that factor in the collision. These categories were chosen because they represent preventable human actions or conditions that are highly correlated with the likelihood and severity of traffic incidents. (see @tbl-cleaned-human-factors-data)

1. **speeding**: Refers to collisions where speeding was a contributing factor. Speeding reduces a driver's ability to react quickly to road conditions or obstacles, and it increases the severity of crashes when they occur, making it one of the primary causes of traffic collisions.

2. **aggressive_and_disturbed_driving**: This category examines whether aggressive behavior, such as reckless driving, or distracted driving (like texting or talking on the phone) was involved in the collision. Aggressive and distracted driving are known to impair judgment and reaction time, leading to accidents that could otherwise have been avoided.

3. **alcohol**: Evaluates if alcohol was a factor in the collision. Alcohol impairs motor skills and decision-making, and is a well-documented contributor to severe and fatal traffic accidents. This category is crucial for understanding the role of substance use in driver behavior.

4. **disability**: Assesses if any medical or physical disability played a role in the collision. Disabilities, whether temporary or permanent, can affect a driver's ability to control the vehicle or respond to changing road conditions, making this an important factor to consider in the context of traffic safety.

These categories were selected because they represent key human factors that have a direct impact on traffic safety, and addressing them can help reduce both the frequency and severity of accidents.

```{r}
#| message: false
#| echo: false
#| label: fig-frequency-factors-tall
#| fig-cap: Frequency of human factors with thinner numbers and taller graph

# Convert relevant columns to factors and reorder them for better readability
human_factors_data <- human_factors_data %>%
  mutate(speeding = factor(speeding, levels = c("None", "Yes")),
         aggressive_and_disturbed_driving = factor(aggressive_and_disturbed_driving, levels = c("None", "Yes")),
         alcohol = factor(alcohol, levels = c("None", "Yes")),
         disability = factor(disability, levels = c("None", "Yes")))

# Frequency summary for each factor
factor_summary <- human_factors_data %>%
  summarise(
    speeding = sum(speeding == "Yes"),
    aggressive_and_disturbed_driving = sum(aggressive_and_disturbed_driving == "Yes"),
    alcohol = sum(alcohol == "Yes"),
    disability = sum(disability == "Yes")
  ) %>%
  tidyr::gather(key = "factor", value = "count")

# Split the category name for 'aggressive_and_disturbed_driving'
factor_summary <- factor_summary %>%
  mutate(factor = str_replace_all(factor, "aggressive_and_disturbed_driving", "aggressive\nand disturbed driving"))

# Reorder factors based on count for better visualization
factor_summary <- factor_summary %>%
  mutate(factor = forcats::fct_reorder(factor, count, .desc = TRUE))

# Bar plot with thinner numbers and taller graph
ggplot(factor_summary, aes(x = factor, y = count, fill = factor)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = count), vjust = -0.5, nudge_y = 200, size = 3, fontface = "plain", color = "black") +  # Thinner numbers
  theme_minimal() +
  labs(title = "Frequency of Human Factors in Traffic Collisions",
       x = "Human Factors",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Adjust x-axis labels
  coord_cartesian(ylim = c(0, 11000))  # Adjust the y-axis limit to make the graph taller

```

```{r}
#| message: false
#| echo: false
#| label: fig-trend-over-time
#| fig-cap: Trends of human factors in traffic collisions over time

# Convert the date column to date format
human_factors_data <- human_factors_data %>%
  mutate(date = as.Date(date, format = "%Y-%m-%d"))

# Summarise the presence of each factor by year
trend_data <- human_factors_data %>%
  mutate(year = lubridate::year(date)) %>%
  group_by(year) %>%
  summarise(
    speeding = sum(speeding == "Yes"),
    aggressive_and_disturbed_driving = sum(aggressive_and_disturbed_driving == "Yes"),
    alcohol = sum(alcohol == "Yes"),
    disability = sum(disability == "Yes")
  ) %>%
  tidyr::gather(key = "factor", value = "count", -year)

# Line plot to show trends over time
ggplot(trend_data, aes(x = year, y = count, color = factor)) +
  geom_line(linewidth = 0.8) +
  theme_minimal() +
  labs(title = "Trends in Human Factors Over Time",
       x = "Year",
       y = "Count of Collisions",
       color = "Human Factor")
```

The second dataset focuses on the environmental factors of traffic collisions, capturing the conditions that surround and influence accidents. These factors play a critical role in how drivers respond to the road and other vehicles, and they can either increase or mitigate the risk of collisions. The selected categories—'traffic_control_type', 'visibility', 'light', and 'road_surface_condition'—were chosen because they reflect the external conditions that can directly affect driver behavior and vehicle performance during a collision. Each category contributes to understanding the environment in which the collision occurred and highlights preventable factors or areas for safety improvement. (see @tbl-cleaned-environmental-factors-data)

1. **traffic_control_type**: This category identifies the type of traffic control present at the scene, such as "Stop Sign," "Traffic Signal," "Pedestrian Crossover," or "No Control." These labels were chosen because the presence and type of traffic control significantly influence driver decisions. Proper traffic control can prevent accidents, while its absence or failure can lead to dangerous situations.

2. **visibility**: This category describes the environmental visibility conditions at the time of the collision, such as "Clear," "Fog, Mist, Smoke, Dust," "Rain," or "Drifting Snow." Visibility is critical for a driver’s ability to react to road hazards or changes. Poor visibility, whether caused by weather or other environmental factors, can lead to misjudgments or delays in reaction time, making it an important factor to analyze in traffic collisions.

3. **Light**: This refers to the lighting conditions at the time of the collision, with possibilities such as "Daylight," "Dusk," "Dark," or artificial lighting conditions like "Dark, artificial" or "Dusk, artificial." Lighting impacts how well a driver can see the road, pedestrians, and other vehicles. Dim or changing light conditions can impair a driver’s perception, increasing the likelihood of an accident, which is why this category is crucial for safety assessments.

4. **road_surface_condition**: This captures the state of the road surface, including conditions like "Dry," "Wet," "Loose Snow," "Ice," or "Slush." The condition of the road surface affects the vehicle’s traction and stability, making it one of the key determinants of how well a driver can maintain control of their vehicle. Poor road surface conditions, such as ice or wet roads, are a known cause of accidents and were chosen for their direct impact on vehicle performance.

These categories were selected because they represent the most critical environmental factors that influence traffic safety. Understanding and addressing these factors can improve road designs, traffic control systems, and driver awareness, ultimately reducing collision rates.

```{r}
#| message: false
#| echo: false
#| label: tbl-cleaned-environmental-factors-data
#| tbl-cap: Sample of Environmental Factors of Collision Data

head (environmental_factors_data, 5) |>
  kable(
    col.names = c("date", "traffic_control_type", "visibility", "light", "road_surface_condition"),
  booktabs = TRUE
  )
```

```{r}
#| message: false
#| echo: false
#| label: fig-four-graphs-layout
#| fig-cap: Combined layout with four simplified collision graphs

# Modify traffic_control_type to group categories into 'No Control', 'Stop Sign', 'Traffic Signal', and 'Others'
environmental_factors_data <- environmental_factors_data %>%
  mutate(traffic_control_type_grouped = case_when(
    traffic_control_type == "No Control" ~ "No Control",
    traffic_control_type == "Stop Sign" ~ "Stop Sign",
    traffic_control_type == "Traffic Signal" ~ "Traffic Signal",
    TRUE ~ "Others"
  ))

# Modify visibility to group categories into 'Clear', 'Rain', 'Snow', and 'Others'
environmental_factors_data <- environmental_factors_data %>%
  mutate(visibility_grouped = case_when(
    visibility == "Clear" ~ "Clear",
    visibility == "Rain" ~ "Rain",
    visibility == "Snow" ~ "Snow",
    TRUE ~ "Others"
  ))

# Common theme
common_theme <- theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),  # Slightly larger font for clarity
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(size = 10),
    plot.margin = margin(t = 15, r = 15, b = 15, l = 15)  # Margins for more space
  )

# First plot: Traffic Control Type and Visibility Conditions

# Plot 1: Traffic Control Type
plot1 <- ggplot(environmental_factors_data, aes(x = traffic_control_type_grouped)) +
  geom_bar(fill = "steelblue") +
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, size = 3) +
  common_theme +
  labs(title = "Collisions by Traffic Control Type",
       x = "Traffic Control Type",
       y = "Number of Collisions") +
  coord_cartesian(ylim = c(0, max(table(environmental_factors_data$traffic_control_type_grouped)) * 1.2))

# Plot 2: Visibility Conditions
plot2 <- ggplot(environmental_factors_data, aes(x = visibility_grouped)) +
  geom_bar(fill = "steelblue") +
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, size = 3) +
  common_theme +
  labs(title = "Collisions by Visibility Conditions",
       x = "Visibility",
       y = "Number of Collisions") +
  coord_cartesian(ylim = c(0, max(table(environmental_factors_data$visibility_grouped)) * 1.2))

# Arrange these two plots side by side
first_plot_pair <- grid.arrange(plot1, plot2, ncol = 2, 
                                top = textGrob("Environmental Factors Affecting Collisions", gp = gpar(fontsize = 15, fontface = "bold")))

```

```{r}
#| message: false
#| echo: false
#| label: tbl-cleaned-environmental-factors-data-2
#| tbl-cap: Sample of Environmental Factors of Collision Data 2

# Common theme
common_theme <- theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    plot.title = element_text(size = 10),  # No bold here
    plot.margin = margin(t = 15, r = 15, b = 15, l = 15)  # Margins for more space
  )

# Modify light to group categories into 'Dark', 'Dawn', 'Daylight', 'Dusk', and 'Others'
environmental_factors_data <- environmental_factors_data %>%
  mutate(lighting_grouped = case_when(
    light %in% c("Dark", "Dark, artificial") ~ "Dark",
    light %in% c("Dawn", "Dawn, artificial") ~ "Dawn",
    light %in% c("Daylight", "Daylight, artificial") ~ "Daylight",
    light %in% c("Dusk", "Dusk, artificial") ~ "Dusk",
    light %in% c("None", "Other") ~ "Others"
  ))

# Modify road_surface_condition to group categories into 'Dry', 'Wet', and 'Other'
environmental_factors_data <- environmental_factors_data %>%
  mutate(road_surface_grouped = case_when(
    road_surface_condition == "Dry" ~ "Dry",
    road_surface_condition == "Wet" ~ "Wet",
    TRUE ~ "Other"
  ))

# Second plot: Lighting Conditions and Road Surface Condition

# Plot 3: Lighting Conditions
plot3 <- ggplot(environmental_factors_data, aes(x = lighting_grouped)) +
  geom_bar(fill = "steelblue") +
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, size = 3) +
  common_theme +  # Apply the common theme
  labs(title = "Collisions by Lighting Conditions",
       x = "Lighting Conditions",
       y = "Collisions") +
  coord_cartesian(ylim = c(0, max(table(environmental_factors_data$lighting_grouped)) * 1.2))

# Plot 4: Road Surface Conditions
plot4 <- ggplot(environmental_factors_data, aes(x = road_surface_grouped)) +
  geom_bar(fill = "steelblue") +
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, size = 3) +
  common_theme +  # Apply the common theme
  labs(title = "Collisions by Road Surface Condition",
       x = "Road Surface Condition",
       y = "Collisions") +
  coord_cartesian(ylim = c(0, max(table(environmental_factors_data$road_surface_grouped)) * 1.2))

# Arrange these two plots side by side
second_plot_pair <- grid.arrange(plot3, plot4, ncol = 2, 
                                 top = textGrob("Environmental Factors Affecting Collisions: Part 2", gp = gpar(fontsize = 15, fontface = "bold")))

```

The last dataset focuses on the result data, specifically analyzing the outcomes of traffic collisions in terms of the 'classification_of_accident' and the 'injury'. These categories were selected because they offer essential insights into the real-world consequences of traffic accidents, both in terms of human harm and material damage. By understanding the outcomes, safety measures and interventions can be better designed to prevent severe accidents in the future. (see @tbl-cleaned-result-data)

1. **classification_of_accident**: This category outlines the type of accident that occurred, with classifications such as "Fatal," "Non-Fatal Injury," "None," and "Property Damage Only." These labels were selected because they provide a clear and structured way to assess the severity of the collision. This classification helps to categorize accidents based on their impact, whether they resulted in fatalities, injuries, or just property damage. By understanding the distribution of these classifications, authorities can prioritize safety interventions to reduce fatal or injury-related accidents.

2. **injury**: This category refers to the level of injury sustained in the accident, with possible values including "Fatal," "Major," "Minor," "Minimal," and "None." Injury severity is a key indicator of the human toll of traffic collisions. These specific levels were selected because they cover the full spectrum of injury outcomes, from minor incidents to fatal accidents. This allows for a detailed understanding of the physical impact on those involved in the accidents, facilitating better emergency response planning and public health interventions.

These categories were chosen because they provide crucial information on the aftermath of traffic collisions, helping to identify patterns in accident severity and injury. Understanding these outcomes is vital for designing targeted measures to reduce the occurrence and impact of serious traffic accidents.

```{r}
#| message: false
#| echo: false
#| label: tbl-cleaned-result-data
#| tbl-cap: Sample of Result of Collision Data

head (result_data, 5) |>
  kable(
    col.names = c("date", "classification_of_accident", "injury"),
  booktabs = TRUE
  )
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References